Open Data for _Crime and Place_ Research: A Practical Guide in R.

### 2020-11-19 manuscript.Rmd
revisions update

Open data initiatives provide a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements, and negotiations [@manovich2011trending]. Open data can lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role]. As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyse open datasets.

This chapter will outline how to develop such skills by providing a framework to approach and meaningfully interpret open data. The supplementary material also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyse different sources of open data in order to draw conclusions about crime and place.

@@

Relatedly, researchers should be aware of issues over sampling and selection, and ultimately, the generalizability of findings that emerge from the analysis of open data. In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduces a degree of bias [@longley2012geodemographics]. Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good]. Contributions to resources such as Open Street Map are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting]. These are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.

@@

- **Where** does this data come from? 
- **Why** was the data collected in the first place?
- **Who** is represented in the data, and who is excluded?
- **What** concepts and constructs can and cannot be operationalized with this data?
- **When** did data collection take place, and how might have that influenced results?

@@

With this in mind, we encourage readers to complete the practical exercise in which we will utilize multiple sources of open data to explore police-recorded crime on and around public transport in London, England.

@@

This chapter has sought to introduce open data as a novel, growing and invaluable tool in crime and place research through a review of key fundamentals and a substantive demonstration within R. We have defined open data, and outlined some key types of open information available, along with their respective strengths and weaknesses. An important component of this review has been to encourage an engaged and critical research approach to using open data. Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources. The supplementary material contained a practical exercise in which we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper. Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line. The demonstration highlighted both the power of open data and the need to critically engage with sources.

@@

As it stands, open data represents a key resource in crime and place research. It is a fundamental component in the movement towards transparent and reproducible scientific research. The skills required to effectively access and use open data, such as those demonstrated in this chapter, are increasingly being taught and deployed in social science. The number and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology. For instance, information about building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime concentrations in the capital, unrestricted by fees or strict terms, for the first time. But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go. As with any emerging domain, open data sources vary considerably in their degree of completeness and reliability. Whilst this can offer insight, for instance, in terms of public participation and sense of community, it can also represent a significant obstacle to empirical analysis. A key challenge for the future will be to ensure that students and research practitioners ask critical questions of their data before blindly delving into analysis and public dissemination. Moreover, the longevity of open data licences remains an unknown parameter. With so much open data being distributed by the public sector, the accessibility of open information is subject to fluctuations in the amount of resource (and goodwill) available to safe-guard continuity. One way of justifying the continued collection and distribution of open public sector data is simply to use the data for public good, such as informing interventions which help reduce crime victimization.

>>>>

Open data initiatives provide a response to these shortcomings, broadening access and participation in research, removing the requirement for permissions, formal agreements, and negotiations [@manovich2011trending]. Open data can lead to all sorts of novel insight within crime and place research, tapping into constructs and processes which are difficult to capture through surveys, interviews and other traditional measures [@solymosi2018role]. As such, it is important that social scientists, researchers, crime analysts, and others interested in making sense of the world around them have the skills and know-how to access, interpret, critique, and analyze open datasets.

This chapter will outline how to develop such skills by providing a framework to approach and meaningfully interpret open data. The supplementary material also offers a practical hands-on guide to demonstrate how to access, wrangle, and analyze different sources of open data in order to draw conclusions about crime and place.

@@

Relatedly, researchers should be aware of issues over sampling and selection, and ultimately, the generalizability of findings that emerge from the analysis of open data. In the case of crowdsourced data, the sample is not randomly drawn from a population, but rather, it is self-selected, giving way for people willing to discuss or contribute to a particular issue, which introduces a degree of bias [@longley2012geodemographics]. Specifically, contributors tend to be men, between the ages of 20-50, with a college or university degree [@budhathoki2010participants; @haklay2010good]. Contributions to resources such as Open Street Map are correlated with contextual characteristics such as poverty and population density, and as such, coverage is non-uniformly distributed across urban areas [@mashhadi2013putting]. These issues are important to keep in mind (and be transparent about) when reporting findings based on analysis of such data.

@@

- **Where** does this data come from? 
- **Why** was the data collected in the first place?
- **Who** is represented in the data, and who is excluded?
- **What** concepts and constructs can and cannot be operationalized with this data?
- **When** did data collection take place, and how might that have influenced results?

@@

With this in mind, we encourage readers to complete the practical exercise in which we utilize multiple sources of open data to explore police-recorded crime in and around public transport stations in London, England.

@@

This chapter has sought to introduce open data as a novel, growing and invaluable tool in crime and place research through a review of key fundamentals. The corresponding supplementary material also includes a substantive demonstration using R. We have defined open data, and outlined some key types of open information available, along with their respective strengths and weaknesses. An important component of this review has been to encourage an engaged and critical research approach to using open data. Open data is both a rich resource of dependent and independent variables for researching the geography of crime, and an interesting research topic in and of itself as a tool for providing insight overlooked by traditional data sources. The supplementary material contained a practical exercise in which we accessed different sources of open data in order to explore crime concentrations at London Underground stations on the Jubilee line: police-recorded crime data via a direct download from a public sector website, transport data via directly querying a public sector API, hosted by Transport for London, and the equivalent transport data retrieved from a crowdsourced database, Open Street Map, using an API wrapper. Minor discrepancies in the geographic location of underground stations in each open data source painted two starkly different stories about the volume and concentration of crime occurring on the line. The demonstration highlighted both the power of open data and the need to critically engage with sources.

@@

As it stands, open data represents a key resource in crime and place research. It is a fundamental component in the movement towards transparent and reproducible scientific research. The skills required to effectively access and use open data, such as those demonstrated in the practical exercise, are increasingly being taught and deployed in social science. The scale and quality of open data resources is also improving, many of which will help illuminate key topics in spatial criminology. Open Street Map represents an ever-growing database to extract theoretically relevant information on the built environment (e.g. land-use, space syntax) and the opportunity structures of micro-places (e.g. locations of bars, gas stations) to study crime and place. New resources are also emerging. For instance, information about historical building construction in Greater London is currently being collected through a crowdsourced platform, _Colouring London_ [@hudson2018colouring], opening prospect for studies to investigate historical urban development and crime in the capital. But, as the discussions and demonstrations in this chapter have shown, there is still a long way to go. As with many domains, open data sources vary considerably in their degree of completeness, validity and reliability. Whilst this can offer insight, for instance, in terms of public participation and sense of community, it can also represent a significant obstacle to empirical analysis. A key challenge for the future will be to ensure that students and research practitioners ask critical questions of their data before blindly delving into analysis and public dissemination. Moreover, the longevity of open data licences remains an unknown parameter. With so much open data being distributed by the public sector, the accessibility of open information is subject to fluctuations in the amount of resource (and goodwill) available to safe-guard continuity. One way of justifying the continued collection and distribution of open public sector data is simply to use the data for public good, such as informing interventions which help reduce crime victimization.

### 2020-11-19 supplementary_materials.Rmd
revisions update

Open Street Map has two main types of wrappers available for its API, a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/), and an R package called `osmdata`.

@@

The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the **Console**.

@@

```{r mapline, eval = F}
ggplot() +
  geom_sf(data = tfl_jub_sf) +
  geom_sf(data = osm_jub_sf, col = "red", alpha = 0.5) 
```

@@

# Create a mini study area for our visualization, by getting the bounding box for the
# Open Street Map (or TfL) buffer. Note that we transform to WGS 84 first in preperation
# for the Open Street Map API.
bb_sf <- stratford_osm_sf %>%
  st_transform(crs = 4326) %>% 
  st_bbox()

# Retrieve the coordinates of the bounding box, and create a numeric list for use with
# osmdata package.
strat_bb <- c(bb_sf[[1]], bb_sf[[2]], bb_sf[[3]], bb_sf[[4]])

@@

Engaging with the creators of the data would also be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed at the beginning of this chapter. For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations. For Open Street Map, its online transparency means that we could find out which user(s) recorded the Jubilee station locations, and directly contact this community of volunteers to find out how station locations were recorded and why.

@@

A useful resource going forward is the _Open Data Handbook_ which can be found online for free (https://opendatahandbook.org/). It provides introductory information (e.g. definitions), advice (e.g. legal, technical details), case studies which showcase the societal value of open data, and an e-library of papers, presentations and videos on open data. It is one of many projects organized by the _Open Knowledge Foundation_. Another comprehensive and free online resource is the _Open Data Institute_ website (https://theodi.org/). Their mission is to demonstrate the positive value of open data and advocate for its ethical creation and usage by working with companies and governments around the globe. For resources geared towards R, _rOpenSci_ (https://ropensci.org/) develop packages and offer an infrastructure and community to promote the accessible usage of R for open scientific research. The official _RStudio_ webpage (https://rstudio.com/about/) also provides a wealth of information and freely available resources for making the most of the software.

>>>>

Two main types of wrappers available for the Open Street Map API are a web-based GUI called Overpass Turbo (https://overpass-turbo.eu/) and an R package called `osmdata`.

@@

The resulting object `osm_stat_sf` contains lots of information. We can view the contents of the object by simply executing the object name into the R Console.

@@

```{r mapline, eval = F}
ggplot() +
  geom_sf(data = osm_jub_sf) +
  geom_sf(data = tfl_jub_sf, color = "red", alpha = 0.5)  
```

@@

# Create a mini study area for our visualization, by getting the bounding box for the
# Open Street Map (or TfL) buffer. Note that we transform to WGS 84 first in
# preperation for the Open Street Map API.
bb_sf <- stratford_osm_sf %>%
  st_transform(crs = 4326) %>% 
  st_bbox()

# Retrieve the coordinates of the bounding box, and create a numeric list for use
# with osmdata package.
strat_bb <- c(bb_sf[[1]], bb_sf[[2]], bb_sf[[3]], bb_sf[[4]])

@@

Engaging with the creators of the data would also be an important step in answering the critical questions ('Where', 'Why', 'Who', 'What' and 'When') posed in the book chapter. For TfL, this might mean investigating the criteria used by their analysts to define the coordinates for stations. For Open Street Map, its online transparency means that we could find out which user(s) recorded the Jubilee station locations, and directly contact this community of volunteers to find out how station locations were recorded and why.

@@

A useful resource going forward is the _Open Data Handbook_ which can be found online for free (https://opendatahandbook.org/). It provides introductory information (e.g. definitions), advice (e.g. legal, technical details), case studies which showcase the societal value of open data, and an e-library of papers, presentations and videos on open data. It is one of many projects organized by the Open Knowledge Foundation. Another comprehensive and free online resource is the _Open Data Institute_ website (https://theodi.org/). Their mission is to demonstrate the positive value of open data and advocate for its ethical creation and usage by working with companies and governments around the globe. For resources geared towards R, _rOpenSci_ (https://ropensci.org/) develop packages and offer an infrastructure and community to promote the accessible usage of R for open scientific research. The official _RStudio_ webpage (https://rstudio.com/about/) also provides a wealth of information and freely available resources for making the most of the software.
